One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction

### 2017-10-18 draft_02.tex
minor changes

\documentclass[aps,pre,twocolumn,groupedaddress]{revtex4-1}
%\documentclass[a4paper,onecolumn,12pt]{article}
%\documentclass[reprint, amsmath,amssymb,aps,]{revtex4-1}

@@

\begin{abstract}
\end{abstract}
 \maketitle
% \tableofcontents
% \newpage
\section{Introduction}
Controlled and tunable experimental realizations of confined quantum systems with ultracold atoms enable nonequilibrium studies of quantum many-body states in novel geometries. A common feature of many cold-atom experiments is the confinement of a many-body system in a harmonic trap. Trapping introduces many distinctive features which have no analog in uniform many-body states, including collective excitations such as breathing modes, dipole modes, and scissors modes. Such trap-related collective modes have been widely studied both experimentally and theoretically for continuum systems, especially in the mean-field regime, since the early days after quantum degeneracy was achieved with trapped atoms \cite{Dalfovo1997}\cite{Jin1996}.

In mean-field perspective, the physics of atom gas in Bose-Einstein Condensation can be described by the Gross-Pitaevskii equation\cite{Dalfovo1999}\cite{Stringari1996}, where the breathing frequency for 1D Bose gas is calculated to be $\Omega_{GP}=\sqrt{3}\omega_0$. Meanwhile, the experiment of Ref.\cite{Haller2009} has found a regime of interactions where the breathing-mode frequency approaches the Gross-Pitaevskii prediction. However, since mean-field approach which generally omit quantum fluctuation is successful in this problem, one may wonder how quantum effects play the role in the breathing mode. Thus, a natural question would be asked is that how breathing frequency is dependent on interaction for classical gas. Though generated in BEC, the breathing mode phenomenon is not restricted to BEC. For a classical gas, the breathing mode still can be excited with a quench. Usually, a good knowledge of the classical version may help us understand the quantum one better. David Guéry-Odelin, Francesca Zambelli, Jean Dalibard, and Sandro Stringari answered this question in ref.\cite{Guery-Odelin1999}. They used the Boltzmann equation as well as molecular dynamics simulation to study the monople mode and quarulple mode of the 2-D classical gas in a harmonic trap. But in that work, the interaction is elastic collision, i.e. zero-range-infinite-strength interaction. We think introducing a finite range for the repulsive interaction will definitely enrich the physics in this system which looks simple, thus producing a different equilibrium and non-equilibrium behaviour. Surprisingly, to the best of our knowledge, there is no in-depth study of this case. For this reason, more study is necessary.

On the other hand, in a nonquilibrium many-body system, a more intriguing question is about its thermalization or relaxation property. Thermalization is among the most fundamental processes for a many-body interacting system. All systems do not thermalize. An example is the Fermi-Pasta-Ulam paradox [ ] which shows confinement in phase-space so that the system stay non-thermalized for a long time. A lot of research on thermalization property has been done in other classical system. Ref.\cite{Tsuchiya2000}\cite{Yawn1997} studied the one dimensional gravitational system with Lyapunov exponents. Ref.\cite{Jin2013} studied the ring of harmonic oscillators and magnetic moment by examine the canonical distribution of a subsystem . In our paper, we will also probe the system's thermalization time scales from the perspective of canonical distribution as well as Lyapunov exponent. %With these approaches, we will show that the thermalization time scale manifest a competition between total energy and interaction energy, so there is not any singular behavior in the thermalization time scales except the trivial divergence at zero interaction.

@@

\subsubsection{Ground states}
For the ground state of this system, one could make the following prediction: \\(1) When the $\sigma$ is larger than the size of the cloud (in view of the scaling behavior, large $\sigma$ is equivalent to small $F_0$ limit when we fix $\sigma=1$), it can be easily shown that the ground state is a solid-like state, where each particle keeps an equal distance ($\frac{2F_0}{\omega_0^2}$) with its nearest neighbors . The low-lying excitation is particles' independent harmonic oscillation around their equilibrium position.\\(2) When $F_0$ is very large the molecules act like a hardcore gas, while the size of particle is just $\sigma$.

@@

\section{Breathing Frequency}\label{section:breathing frequency}
Usually, to see the response of quantum gas to a quench, one begin simulation with the ground state and then suddenly vary some parameter by a little, for instance, $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$. In that case, one will observe the radius R of the cloud oscillating at a certain frequency, which is usually called ``breathing mode".
However, it is quite different in the classical version. Firstly, it is no longer necessary to focus on ground state. In quantum case, the reason why we care about ground state and low lying excitations is that BEC occur at low temperature. But in classical case, there is no BEC. So it is not necessary to focus on low lying excitations in classical case. Besides, since classical limit is $\hbar\rightarrow 0$, no matter how low the kT is, $\hbar/kT$ is always zero so even in low energy regime, the classical behaviors could be quite different from the quantum one. Secondly, it is no longer necessary to vary parameters ``by a little" in the quench. In quantum case, the small quench $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$ means the spectrum of Hamiltonian is shifted slightly, thus one may expect to see some beat phenomenon({\color{red}{?}}). But in classical case, since $\hbar$ vanishes, no matter how small the quench is, the reason that cause beats in quantum case no longer work in classical context. Thus it is not necessary to limit ourselves to small quench.

@@

\bibliography{mybib}

>>>>

\documentclass[aps,preprintnumbers,onecolumn,amsmath,amssymb,floatfix,pra]{revtex4-1}

%\documentclass[aps,preprintnumbers,onecolumn,amsmath,amssymb,floatfix,pra]{revtex4}

%\documentclass[aps,pre,twocolumn,groupedaddress]{revtex4-1}
%\documentclass[a4paper,onecolumn,12pt]{article}
%\documentclass[reprint, amsmath,amssymb,aps,]{revtex4-1}

@@

\begin{abstract}

Abstract --  to be written last

\end{abstract}

\maketitle

\section{Introduction}

Controlled and tunable experimental realizations of confined quantum systems with ultracold atoms
enable nonequilibrium studies of quantum many-body states in novel geometries. A common feature of
many cold-atom experiments is the confinement of a many-body system in a harmonic trap. Trapping
introduces many distinctive features which have no analog in uniform many-body states, including
collective excitations such as breathing modes, dipole modes, and scissors modes. Such trap-related
collective modes have been widely studied both experimentally and theoretically for continuum
systems, especially in the mean-field regime, since the early days after quantum degeneracy was
achieved with trapped atoms \cite{Dalfovo1997}\cite{Jin1996}.

In mean-field perspective, the physics of atom gas in Bose-Einstein Condensation can be described by
the Gross-Pitaevskii equation\cite{Dalfovo1999}\cite{Stringari1996}, where the breathing frequency
for 1D Bose gas is calculated to be $\Omega_{GP}=\sqrt{3}\omega_0$. Meanwhile, the experiment of
Ref.\cite{Haller2009} has found a regime of interactions where the breathing-mode frequency
approaches the Gross-Pitaevskii prediction. However, since mean-field approach which generally omit
quantum fluctuation is successful in this problem, one may wonder how quantum effects play the role
in the breathing mode. Thus, a natural question would be asked is that how breathing frequency is
dependent on interaction for classical gas. Though generated in BEC, the breathing mode phenomenon
is not restricted to BEC. For a classical gas, the breathing mode still can be excited with a
quench. Usually, a good knowledge of the classical version may help us understand the quantum one
better. David Guery-Odelin, Francesca Zambelli, Jean Dalibard, and Sandro Stringari answered this
question in ref.\cite{Guery-Odelin1999}. They used the Boltzmann equation as well as molecular
dynamics simulation to study the monople mode and quarulple mode of the 2-D classical gas in a
harmonic trap. But in that work, the interaction is elastic collision,
i.e. zero-range-infinite-strength interaction. We think introducing a finite range for the repulsive
interaction will definitely enrich the physics in this system which looks simple, thus producing a
different equilibrium and non-equilibrium behaviour. Surprisingly, to the best of our knowledge,
there is no in-depth study of this case. For this reason, more study is necessary.

On the other hand, in a nonquilibrium many-body system, a more intriguing question is about its
thermalization or relaxation property. Thermalization is among the most fundamental processes for a
many-body interacting system. All systems do not thermalize. An example is the Fermi-Pasta-Ulam
paradox [ ] which shows confinement in phase-space so that the system stay non-thermalized for a
long time. A lot of research on thermalization property has been done in other classical
system. Ref.\cite{Tsuchiya2000}\cite{Yawn1997} studied the one dimensional gravitational system with
Lyapunov exponents. Ref.\cite{Jin2013} studied the ring of harmonic oscillators and magnetic moment
by examine the canonical distribution of a subsystem . In our paper, we will also probe the system's
thermalization time scales from the perspective of canonical distribution as well as Lyapunov
exponent. %With these approaches, we will show that the thermalization time scale manifest a
          %competition between total energy and interaction energy, so there is not any singular
          %behavior in the thermalization time scales except the trivial divergence at zero
          %interaction.

@@

\subsubsection{Ground states}

For the ground state of this system, one could make the following prediction: \\(1) When the
$\sigma$ is larger than the size of the cloud (in view of the scaling behavior, large $\sigma$ is
equivalent to small $F_0$ limit when we fix $\sigma=1$), it can be easily shown that the ground
state is a solid-like state, where each particle keeps an equal distance ($\frac{2F_0}{\omega_0^2}$)
with its nearest neighbors . The low-lying excitation is particles' independent harmonic oscillation
around their equilibrium position.\\(2) When $F_0$ is very large the molecules act like a hardcore
gas, while the size of particle is just $\sigma$.

@@

\section{Breathing Frequency}\label{section:breathing frequency}

Usually, to see the response of quantum gas to a quench, one begin simulation with the ground state
and then suddenly vary some parameter by a little, for instance,
$\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$. In that case, one will observe the radius R
of the cloud oscillating at a certain frequency, which is usually called ``breathing mode".

However, it is quite different in the classical version. Firstly, it is no longer necessary to focus
on ground state. In quantum case, the reason why we care about ground state and low lying
excitations is that BEC occur at low temperature. But in classical case, there is no BEC. So it is
not necessary to focus on low lying excitations in classical case. Besides, since classical limit is
$\hbar\rightarrow 0$, no matter how low the kT is, $\hbar/kT$ is always zero so even in low energy
regime, the classical behaviors could be quite different from the quantum one. Secondly, it is no
longer necessary to vary parameters ``by a little" in the quench. In quantum case, the small quench
$\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$ means the spectrum of Hamiltonian is shifted
slightly, thus one may expect to see some beat phenomenon({\color{red}{?}}). But in classical case,
since $\hbar$ vanishes, no matter how small the quench is, the reason that cause beats in quantum
case no longer work in classical context. Thus it is not necessary to limit ourselves to small
quench.

@@

%\bibliography{mybib}
%\bibliographystyle{ieeetr}

\begin{thebibliography}{1}

\bibitem{Dalfovo1997}
F.~Dalfovo, S.~Giorgini, M.~Guilleumas, L.~Pitaevskii, and S.~Stringari,
  ``Collective and single-particle excitations of a trapped bose gas,'' {\em
  Phys. Rev. A}, vol.~56, pp.~3840--3845, Nov. 1997.

\bibitem{Jin1996}
D.~S. Jin, J.~R. Ensher, M.~R. Matthews, C.~E. Wieman, and E.~A. Cornell,
  ``Collective excitations of a bose-einstein condensate in a dilute gas,''
  {\em Phys. Rev. Lett.}, vol.~77, pp.~420--423, July 1996.

\bibitem{Dalfovo1999}
F.~Dalfovo, S.~Giorgini, L.~P. Pitaevskii, and S.~Stringari, ``Theory of
  bose-einstein condensation in trapped gases,'' {\em Rev. Mod. Phys.},
  vol.~71, pp.~463--512, Apr. 1999.

\bibitem{Stringari1996}
S.~Stringari, ``Collective excitations of a trapped bose-condensed gas,'' {\em
  Phys. Rev. Lett.}, vol.~77, pp.~2360--2363, Sept. 1996.

\bibitem{Haller2009}
E.~Haller, M.~Gustavsson, M.~J. Mark, J.~G. Danzl, R.~Hart, G.~Pupillo, and
  H.-C. Nägerl, ``Realization of an excited, strongly correlated quantum gas
  phase,'' {\em Science}, vol.~325, p.~1224, Sept. 2009.

\bibitem{Guery-Odelin1999}
D.~Gu\'ery-Odelin, F.~Zambelli, J.~Dalibard, and S.~Stringari, ``Collective
  oscillations of a classical gas confined in harmonic traps,'' {\em Phys. Rev.
  A}, vol.~60, pp.~4851--4856, Dec 1999.

\bibitem{Tsuchiya2000}
T.~Tsuchiya and N.~Gouda, ``Relaxation and lyapunov time scales in a
  one-dimensional gravitating sheet system,'' {\em Phys. Rev. E}, vol.~61,
  pp.~948--951, Jan. 2000.

\bibitem{Yawn1997}
K.~R. Yawn and B.~N. Miller, ``Ergodic properties and equilibrium of
  one-dimensional self-gravitating systems,'' {\em Phys. Rev. E}, vol.~56,
  pp.~2429--2436, Sept. 1997.

\bibitem{Jin2013}
F.~Jin, T.~Neuhaus, K.~Michielsen, S.~Miyashita, M.~A. Novotny, M.~I.
  Katsnelson, and H.~D. Raedt, ``Equilibration and thermalization of classical
  systems,'' {\em New Journal of Physics}, vol.~15, no.~3, p.~033009, 2013.

\end{thebibliography}
